---
title: "Altercentric pupils - pilot"
author: LL
output: pdf_document
date: `11/04/2025`
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
#rm(list=c("try"))
#rm(list=ls())
library(tidyverse)
library(arrow)
library(tidyr)
library(stringr)
library(purrr)
library(summarytools)
library(exactRankTests)
library(lme4)
library(naniar)
#library(devtools)
remotes::install_github("dmirman/gazer")
library(gazer)
library(zoo)
library(wesanderson)
library(gcookbook)
source("functions/diagnostic_fcns.r")
source("functions/glmm_stability.r")
source("functions/boot_glmm.r")


library(itsadug)
packageVersion("itsadug")
library(plotfunctions)
packageVersion("plotfunctions")
library(colorspace)
packageVersion("colorspace")
## Define colors:
col1 <- 'pink1'
col2 <- 'black'
col3 <- 'indianred'
```


## Load data

```{r}
## time.frame for interpolation
max.time <- 15367 #occ act
min.time <- 0
time <- seq(from = min.time, to = max.time, by = 1)
xx <- as.data.frame(time)
str(xx)

baseline.end<-8517 #occ act
#frozen frame with actress looking at unoccluded platform (from both POVs) with or without object, after hand has disappeared (whole video IP time, in ms)

#this will be counterbalancing file
# demo.data <- read_csv("experiment_2/data/Woodward_2022_counterbalancing.csv") %>%
#   mutate(
#     EDF1 = fct_recode(
#       as.factor(EDF1),
#       "Georgia_1" = "Georgia1",
#     )
#   ) %>%
#   separate(EDF1, c("subject", "num"), sep = "_")%>%
#   select(-num)
```

```{r}
#convert sample data to parquet
#sample reports saved on Toshiba drive

wd <- "D:/altercentric_pupil/DataViewingSession_altercentric_pupils_infants/Output"
xdata<-read.table(file.path(wd, "sample_report_whole_video.txt"), sep ="\t", header = TRUE) 
str(xdata)

xdata2<-xdata %>%
  select(RECORDING_SESSION_LABEL, TRIAL_INDEX, IP_LABEL, IP_START_TIME, IP_DURATION, HTARGET_DISTANCE, RIGHT_GAZE_X, RIGHT_GAZE_Y, RIGHT_INTEREST_AREAS, RIGHT_INTEREST_AREA_DATA, RIGHT_INTEREST_AREA_ID, RIGHT_INTEREST_AREA_LABEL, RIGHT_INTEREST_AREA_PIXEL_AREA, RIGHT_IN_BLINK, RIGHT_IN_SACCADE, RIGHT_PUPIL_SIZE, SAMPLE_INDEX, SAMPLE_MESSAGE, SAMPLING_RATE, TIMESTAMP, TRIAL_START_TIME, VIDEO_FRAME_INDEX, fixation_trigger, sound_file, Trial_Index_, trial_number, condition, counterb_id, video_file, duration_video, bg_name)

write_parquet(xdata2, "sample_report_whole_video.parquet")

sample.data<-read_parquet("./data/sample_report_whole_video.parquet")
```

load sample data and add the demographic data
```{r}
sample.data<-read_parquet("./data/sample_report_whole_video.parquet") %>%
  #exclude trials that were not shown due to experimenters interrupting the experiment 
  #(due to malfunction or fussiness)
  filter(!(RECORDING_SESSION_LABEL == "pilot2" & trial_number == 2), 
         !(RECORDING_SESSION_LABEL == "pilot4" & trial_number == 17),
         !(RECORDING_SESSION_LABEL == "pilot6" & trial_number == 9),
         !(RECORDING_SESSION_LABEL == "pilot7" & trial_number == 5),
         !(RECORDING_SESSION_LABEL == "pilot14" & trial_number == 3),
         !(RECORDING_SESSION_LABEL == "pilot16" & trial_number == 14),
         !(RECORDING_SESSION_LABEL == "pilot18" & trial_number == 6)) %>% 
  mutate(condition_trial = paste(condition, "_", Trial_Index_)) %>%
  filter(trial_number != 11)%>% #skip cartoon
 # full_join(demo.data)%>%
  mutate(time = TIMESTAMP - IP_START_TIME) %>% 
  mutate( # Remove .mp4 and split by underscore
    name_no_ext = str_remove(video_file, "\\.mp4$"),
    parts = str_split(name_no_ext, "_")
  ) %>%
  mutate(
    perspective = map_chr(parts, ~ paste(.x[1:2], collapse = "_")),
    toy = map_chr(parts, ~ .x[3]),
    object_presence = map_chr(parts, ~ .x[4]),
    agent_side = map_chr(parts, ~ ifelse("mirrored" %in% .x, "right", "left")),
    
    # Replace shorthand values with full labels
    perspective = case_when(
      perspective == "occ_act" ~ "occlusion_from_actor",
      perspective == "occ_part" ~ "occlusion_from_participant",
      perspective == "occ_neith" ~ "occlusion_from_neither",
      TRUE ~ perspective
    ),
    object_presence = case_when(
      object_presence == "abs" ~ "absent",
      object_presence == "pres" ~ "present",
      TRUE ~ object_presence
    )
  ) %>%
  select(-name_no_ext, -parts)
```


```{r}
#data checks and preparation
str(sample.data)
levels(as.factor(sample.data$counterb_id))
table(sample.data$RECORDING_SESSION_LABEL,sample.data$trial_number)

table(sample.data$RECORDING_SESSION_LABEL,sample.data$condition)
table(sample.data$counterb_id,sample.data$agent_side)
table(sample.data$counterb_id,sample.data$toy)
summary(sample.data$time)
summary(sample.data$VIDEO_FRAME_INDEX)

sample.data$RIGHT_PUPIL_SIZE<-as.numeric(sample.data$RIGHT_PUPIL_SIZE)
summary(sample.data$RIGHT_PUPIL_SIZE)
sample.data$RIGHT_GAZE_X<-as.numeric(sample.data$RIGHT_GAZE_X)
summary(sample.data$RIGHT_GAZE_X)
sample.data$RIGHT_GAZE_Y<-as.numeric(sample.data$RIGHT_GAZE_Y)
summary(sample.data$RIGHT_GAZE_Y)
```
Explore pupil size data
```{r}
#set IPs

# baseline
bs.beg.occ.part<-6.5
bs.end.occ.part<-7.5
bs.beg.occ.act<-8.5
bs.end.occ.act<-9.5
bs.beg.occ.neith<-8.5
bs.end.occ.neith<-9.5
  
#object presence manipulation
beg.occ.part<- 3
end.occ.part<- bs.beg.occ.part
beg.occ.act<- 5.5
end.occ.act<- bs.beg.occ.act
beg.occ.neith<- 5
end.occ.neith<- bs.beg.occ.neith
  
baseline.ip.data<-sample.data %>%
mutate(time_sec=time/1000) %>%
  filter(
    (grepl("act", condition)   & time_sec >= bs.beg.occ.act   & time_sec <= bs.end.occ.act) |
    (grepl("part", condition)  & time_sec >= bs.beg.occ.part  & time_sec <= bs.end.occ.part) |
    (grepl("neith", condition) & time_sec >= bs.beg.occ.neith & time_sec <= bs.end.occ.neith)
  )

#check how many trials we would have to exclude if we only keep trials with at least 50% gaze samples tracked during measuring IP (i.e., 2 out of 4 seconds)
baseline.ip.data.trials<-baseline.ip.data %>% 
  group_by(RECORDING_SESSION_LABEL, trial_number) %>% 
  summarise(tot_number_expected_samples=length(time_sec),
            collected_samples=sum(!is.na(RIGHT_PUPIL_SIZE)),
            prop.valid.samples=collected_samples/tot_number_expected_samples)

sum(baseline.ip.data.trials$prop.valid.samples<=0.5)
mean(baseline.ip.data.trials$prop.valid.samples)
```
```{r}
# object manipulation IP
object.ip.data <- sample.data %>%
  mutate(time_sec=time/1000) %>% 
 filter(
    (grepl("act", condition)   & time_sec >= beg.occ.act   & time_sec <= end.occ.act) |
    (grepl("part", condition)  & time_sec >= beg.occ.part  & time_sec <= end.occ.part) |
    (grepl("neith", condition) & time_sec >= beg.occ.neith & time_sec <= end.occ.neith)
  )

#check data loss in specific IPs
object.ip.data$RIGHT_PUPIL_SIZE<-as.numeric(object.ip.data$RIGHT_PUPIL_SIZE)
summary(object.ip.data$RIGHT_PUPIL_SIZE)
30813/length(object.ip.data$RIGHT_PUPIL_SIZE) #prop. NAs over total samples #25%
 
#check how many trials we would have to exclude if we only keep trials with at least 50% gaze samples tracked during measuring IP (i.e., 2 out of 4 seconds)
object.ip.data.trials<-object.ip.data %>% 
  group_by(RECORDING_SESSION_LABEL, trial_number) %>% 
  summarise(tot_number_expected_samples=length(time_sec),
            collected_samples=sum(!is.na(RIGHT_PUPIL_SIZE)),
            prop.valid.samples=collected_samples/tot_number_expected_samples)

sum(object.ip.data.trials$prop.valid.samples<=0.5)#12 trials to be excluded 
12/length(object.ip.data.trials$tot_number_expected_samples) #7%

mean(object.ip.data.trials$prop.valid.samples)
```


```{r}
# measuring IP
measuring.ip.data <- sample.data %>%
  mutate(time_sec=time/1000) %>% 
  group_by(RECORDING_SESSION_LABEL, trial_number) %>%
  mutate(max_time_sec = max(time_sec)) %>%
  filter(time_sec >= max_time_sec - 4 & time_sec <= max_time_sec) %>%
  ungroup() #%>%
  #select(-max_time_sec)  # optional: remove if not needed

#check data loss in specific IPs
measuring.ip.data$RIGHT_PUPIL_SIZE<-as.numeric(measuring.ip.data$RIGHT_PUPIL_SIZE)
summary(measuring.ip.data$RIGHT_PUPIL_SIZE)
 84439/length(measuring.ip.data$RIGHT_PUPIL_SIZE) #prop. NAs over total samples #25%
 
#check how many trials we would have to exclude if we only keep trials with at least 50% gaze samples tracked during measuring IP (i.e., 2 out of 4 seconds)
measuring.ip.data.trials<-measuring.ip.data %>% 
  group_by(RECORDING_SESSION_LABEL, trial_number) %>% 
  summarise(tot_number_expected_samples=length(time_sec),
            collected_samples=sum(!is.na(RIGHT_PUPIL_SIZE)),
            prop.valid.samples=collected_samples/tot_number_expected_samples)

sum(measuring.ip.data.trials$prop.valid.samples<=0.5)#36 samples to be excluded 
36/length(measuring.ip.data.trials$tot_number_expected_samples)

mean(measuring.ip.data.trials$prop.valid.samples)
```

##LL: OLD, adapt frm here. Artefacts check
*Plot raw data
```{r}
raw_plot_pupil<-ggplot(data = sample.data, aes(x = time, y = RIGHT_PUPIL_SIZE)) +
  ylab("Pupil size") +
  xlab("Time (in ms)") +
  geom_point(aes(color = condition), alpha = 0.3, size = 0.5) +
  facet_grid(agent~subject)+
  xlim(0, 22000) +
    theme_bw()+
  scale_color_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("new goal","new side"))+
  scale_fill_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("new goal","new side"))+
  guides(colour = guide_legend(override.aes = list(size=4)))+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.title = element_blank(), legend.position = c(0.9, 0.9), legend.text=element_text(size=12))

raw_plot_pupil

ggsave(raw_plot_pupil, filename = "graphs/raw_plot_pupil.png", width=12, height=8)
```
*Plot raw data only in the IP of the baseline + analysis
```{r}
raw_plot_pupil_IP<-ggplot(data = sample.data, aes(x = time, y = RIGHT_PUPIL_SIZE)) +
  ylab("Pupil size") +
  xlab("Time (in ms)") +
  geom_point(aes(color = condition), alpha = 0.3, size = 0.5) +
  geom_vline(xintercept = 4000,linetype="longdash", colour="gray")+
  geom_vline(xintercept = 3000,linetype="longdash", colour="grey")+
  annotate("rect", xmin = 3000, xmax = 4000, ymin = 0, ymax = 15000,
           alpha = .3,fill = "grey")+
  facet_grid(agent~subject)+
  xlim(3000, 8001) +
  theme_bw()+
  scale_color_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("new goal","new side"))+
  scale_fill_manual(values=wes_palette("Zissou1")[c(1,4)], labels= c("new goal","new side"))+
  guides(colour = guide_legend(override.aes = list(size=4)))+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.title = element_blank(), legend.position = c(0.9, 0.9), legend.text=element_text(size=12),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


raw_plot_pupil_IP

ggsave(raw_plot_pupil_IP, filename = "graphs/raw_plot_pupil_IP.png", width=12, height=8)
```


* Plot with blink artefacts removed
```{r }
sample.data<-sample.data%>%
        mutate(RIGHT_PUPIL_SIZE_no_blinks=extend_blinks(RIGHT_PUPIL_SIZE,  hz=1000, fillback=100, fillforward=100))#Extends blinks


artefact_check <-
  ggplot(data = sample.data,
         aes(x = time, y = RIGHT_PUPIL_SIZE_no_blinks)) +
  ylab("Pupil size") +
  xlab("Time (in ms)") +
  geom_point(aes(color = condition), alpha = 0.3, size = 0.5) +
  facet_grid(agent ~ subject) +
  xlim(0, 22000) +
  theme_bw() +
  scale_color_manual(values = wes_palette("Zissou1")[c(1, 4)],
                     labels = c("new goal", "new side")) +
  scale_fill_manual(values = wes_palette("Zissou1")[c(1, 4)],
                    labels = c("new goal", "new side")) +
  guides(colour = guide_legend(override.aes = list(size=4)))+
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_blank(), legend.position = c(0.9, 0.9), legend.text=element_text(size=14))


artefact_check

ggsave(artefact_check, filename = "graphs/artefact_check_100.png", width=12, height=8, dpi=600)
```

* Plot with blink artefacts removed only in the IP of the baseline + analysis
```{r}
artefact_check_IP <-
  ggplot(data = sample.data,
         aes(x = time, y = RIGHT_PUPIL_SIZE_no_blinks)) +
  ylab("Pupil size") +
  xlab("Time (in ms)") +
  geom_point(aes(color = condition), alpha = 0.3, size = 0.5) +
  geom_vline(xintercept = 4000,linetype="longdash", colour="grey")+
  geom_vline(xintercept = 3000,linetype="longdash", colour="grey")+
  annotate("rect", xmin = 3000, xmax = 4000, ymin = 0, ymax = 15000,
           alpha = .3,fill = "grey")+
  facet_grid(agent ~ subject) +
  xlim(3000, 8001) +
  theme_bw() +
  scale_color_manual(values = wes_palette("Zissou1")[c(1, 4)],
                     labels = c("new goal", "new side")) +
  scale_fill_manual(values = wes_palette("Zissou1")[c(1, 4)],
                    labels = c("new goal", "new side")) +
  guides(colour = guide_legend(override.aes = list(size=4)))+
  theme(
    panel.grid.major.x = element_blank(),
    panel.grid.minor.x = element_blank(),
    panel.grid.major.y = element_blank(),
    panel.grid.minor.y = element_blank(),
    legend.title = element_blank(), legend.position = c(0.9, 0.9), legend.text=element_text(size=14),
     axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))


artefact_check_IP

ggsave(artefact_check_IP, filename = "graphs/artefact_check_IP.png", width=12, height=8, dpi=600)

```


##Proportion of valid data
```{r}
prop.tracked.data<-sample.data%>%
  group_by(subject, trial_number)%>%
  summarise(length=length(RIGHT_PUPIL_SIZE_no_blinks), sum_noNA=sum(!is.na(RIGHT_PUPIL_SIZE_no_blinks)), prop=sum(!is.na(RIGHT_PUPIL_SIZE_no_blinks))/length(RIGHT_PUPIL_SIZE_no_blinks), min(RIGHT_PUPIL_SIZE), max(RIGHT_PUPIL_SIZE))
prop.tracked.data
mean(prop.tracked.data$prop) #mean eye-detected-time
```


```{r}
prop.tracked.data%>%group_by(subject, trial_number)%>%summarise(min(prop))
#this confirms that 2 test trials (Juna session 2 and Maylo session 3) have less than 70% tracked data (46% and 64% respectively)

prop.tracked.gaze.data<-sample.data%>%
  filter(RIGHT_GAZE_X<1024&RIGHT_GAZE_X>0&RIGHT_GAZE_Y<768&RIGHT_GAZE_Y>0)%>%
  group_by(subject, trial_number)%>%
  summarise(prop=sum(!is.na(RIGHT_GAZE_X))/max(time))
prop.tracked.gaze.data
mean(prop.tracked.gaze.data$prop) #mean on-screen-looking time

prop.tracked.gaze.data%>%group_by(subject)%>%summarise(min(prop), mean(prop))
#for Edgar_2 the eye was detected 93% of the trial time but it was only inside the video area for 66% of the time; 
#Joker_1 eye detected:94%; on screen:67%
#Timo_3 eye detected: 98%; on screen:68%

```
* Plot distribution of pupil sizes
```{r}
puphist <-
  ggplot(sample.data
    #  %>% filter(subject != "Juna" | num != "2") %>%
    #                 filter(subject != "Maylo" | num != "3")
    ,
    aes(x = RIGHT_PUPIL_SIZE_no_blinks)) +
    geom_histogram(aes(y = ..count..),
                     colour = "green", binwidth = 0.5)  +
  xlab("Pupil Size") + ylab("Count") + theme_bw()
puphist
```
## Preprocessing: interpolation, baseline correction, down sampling
```{r}
sample.data.no.missing<-sample.data%>% filter(subject != "Juna" | trial_number != "2") %>%
                            filter(subject != "Maylo" | trial_number != "3") #exclude the 2 test trials in which the eye was not tracked for more than 30% of the trial time
#check
table(sample.data.no.missing$subject, sample.data.no.missing$trial_number)
table(sample.data$subject, sample.data$trial_number)


# determine baseline
exp.data.pupil.base <- sample.data %>%
  filter(time < baseline.end & time > (baseline.end-1000)) %>% 
    group_by(subject, trial_number) %>%
  summarise(median.base.pupil = median(RIGHT_PUPIL_SIZE_no_blinks, na.rm = TRUE))

#preprocessing
exp.data.pupil.processed <- sample.data %>%
  filter(time<max.time)%>%
  select(subject, time, trial_number, condition, agent, RIGHT_GAZE_X, RIGHT_GAZE_Y, RIGHT_PUPIL_SIZE_no_blinks) %>%
  group_by(subject, trial_number, time) %>%
  full_join(xx%>%select(time)) %>% #add missing time.frames
  ungroup() %>%
  group_by(subject, trial_number) %>%
  mutate(pupil.inter = na.approx(RIGHT_PUPIL_SIZE_no_blinks, na.rm = FALSE, maxgap = 500)) %>% #linear interpolation
  full_join(exp.data.pupil.base) %>% #add baseline data
  mutate(pupil.base.corrected = pupil.inter - median.base.pupil)%>% #subtractive baseline correction
  ungroup()%>%
    mutate(bin = cut(time, seq(min(time), max(time), 100), right = FALSE))%>% #addition of time bins (100 ms = 10 hz)
  separate(bin, c("bin_low", "bin_high"), sep=",", remove=FALSE)%>%
  select(-bin_high)%>%
  mutate(bin_low=as.numeric(str_replace_all(bin_low, "\\[|\\]", "")))
  
#down sampling to 10hz using median values:
exp.data.pupil.processed.downsampled <- exp.data.pupil.processed %>%  
  group_by(subject, condition, agent, trial_number, bin_low)%>%
  summarise(pupil.base.corrected.binned=median(pupil.base.corrected), Xgaze=median(RIGHT_GAZE_X), Ygaze=median(RIGHT_GAZE_Y))%>%
  filter(!is.na(subject))%>%
  ungroup()
```


* Plot interpolated data: individual level
```{r}
exp.data.pupil.processed.downsampled2 <-
  exp.data.pupil.processed.downsampled %>%
  mutate(condition = fct_recode(condition, "New goal" = "new_goal", "New side" =
                                  "new_side"))

plot.woodward.pupil.individual<-ggplot(data = exp.data.pupil.processed.downsampled2, aes(x = bin_low, y =pupil.base.corrected.binned )) +
  ylab("Pupil size ") +
  xlab("Time (in ms)") +
    geom_vline(aes(xintercept=baseline.end), lty=2, alpha=0.3)+
  geom_point(aes(color = as.factor(condition)), alpha = 0.8, size = 0.5) +
  geom_line(aes(color = as.factor(condition)), alpha = 0.8) +
  facet_grid(agent~subject)+
  xlim(0, 22000) +
    theme_bw()+
    scale_color_manual(values = wes_palette("Zissou1")[c(1, 4)],
                     labels = c("New goal", "New side")) +
  scale_fill_manual(values = wes_palette("Zissou1")[c(1, 4)],
                    labels = c("New goal", "New side")) +
  guides(colour = guide_legend(override.aes = list(size=4)))+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.title = element_blank(), legend.position = c(0.9,0.9),
        legend.key.size = unit(0.2, "cm"))

plot.woodward.pupil.individual
ggsave(plot.woodward.pupil.individual, filename = "graphs/pupil_processed_baseline_corrected_individual.png", height = 7, width = 20, scale = 1)
```
* Plot interpolated data: individual level, only in the IP
```{r}
plot.woodward.pupil.individual_IP<-ggplot(data = exp.data.pupil.processed.downsampled2, aes(x = bin_low, y =pupil.base.corrected.binned )) +
  ylab("Pupil size ") +
  xlab("Time (in ms)") +
  geom_vline(aes(xintercept=baseline.end), lty=2, alpha=0.3)+
  geom_vline(aes(xintercept=3000), lty=2, alpha=0.3)+
  annotate("rect", xmin = 3000, xmax = 4000, ymin = 0, ymax = 15000,
           alpha = .3,fill = "grey")+
  geom_point(aes(color = as.factor(condition)), alpha = 0.8, size = 0.5) +
  geom_line(aes(color = as.factor(condition)), alpha = 0.8) +
  facet_grid(agent~subject)+
  xlim(3000, 8001) +
    theme_bw()+
    scale_color_manual(values = wes_palette("Zissou1")[c(1, 4)],
                     labels = c("New goal", "New side")) +
  scale_fill_manual(values = wes_palette("Zissou1")[c(1, 4)],
                    labels = c("New goal", "New side")) +
  guides(colour = guide_legend(override.aes = list(size=4)))+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.title = element_blank(), legend.position = c(0.9,0.9),
        legend.key.size = unit(0.2, "cm"))

plot.woodward.pupil.individual_IP
ggsave(plot.woodward.pupil.individual, filename = "graphs/pupil_processed_baseline_corrected_individual_IP.png", height = 7, width = 20, scale = 1, dpi=600)
```
# Plot group level data 
```{r}

pupil.group.level <- exp.data.pupil.processed.downsampled2 %>%
  #mutate(TRIAL_LABEL=recode(TRIAL_LABEL, "Trial: 1"="Trial 1", "Trial: 2"="Trial 2"))%>%
  group_by(bin_low, condition, agent) %>%
  summarise(mean.pupil.corrected.binned = mean(pupil.base.corrected.binned, na.rm = TRUE), sd.pupil.corrected.binned= sd(pupil.base.corrected.binned, na.rm = TRUE), se.pupil.corrected.binned = sd(pupil.base.corrected.binned, na.rm = TRUE) / sqrt(length(pupil.base.corrected.binned)))

plot.woodward.pupil.group <- ggplot(data = pupil.group.level, aes(x = bin_low, y = mean.pupil.corrected.binned)) +
  geom_vline(aes(xintercept=baseline.end), lty=2, alpha=0.3)+
  #geom_vline(aes(xintercept=3000), lty=2, alpha=0.3)+
  # annotate("rect", xmin = 3000, xmax = 4000, ymin =0, ymax = 1000,
  #          alpha = .3,fill = "grey")+
     ylab("Pupil size") +
  xlab("Time (in ms)") +
  geom_point(aes(x = bin_low, y = mean.pupil.corrected.binned, color = condition), alpha = 0.9, size = 0.5) +
  geom_ribbon(aes(ymin = mean.pupil.corrected.binned - se.pupil.corrected.binned, ymax = mean.pupil.corrected.binned + se.pupil.corrected.binned, fill = condition), alpha = 0.7) +
  xlim(3000, 8001) +
    theme_bw()+
  scale_color_manual(values = wes_palette("Zissou1")[c(4, 1)],
                     labels = c("old side - new identity", "old identity - new side")) +
  scale_fill_manual(values = wes_palette("Zissou1")[c(4, 1)],
                    labels = c("old side - new identity", "old identity - new side")) +
  guides(colour = guide_legend(override.aes = list(size=2)))+
  theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank(),
        legend.title = element_blank(), 
        #legend.position = "bottom",
        legend.position = c(0.87, 0.92), 
        legend.margin = margin(0.1,0.1,0.01,0.1),
        legend.text=element_text(size=8),
        legend.key.size = unit(0.4, 'cm'),
        axis.text=element_text(size=9),
        axis.title=element_text(size=12))+
  facet_wrap(~agent)

plot.woodward.pupil.group

ggsave(plot.woodward.pupil.group, filename = "graphs/pupil_group_condition3.png", height = 5, width = 10, scale = 0.7, dpi=600)
```

#GAMM
* Plot gaze positions
```{r}
emptyPlot(c(0,1024), c(768, 0), bty='o',
          main="Gaze positions", xlab="Xgaze", ylab="Ygaze")
points(exp.data.pupil.processed.downsampled$Xgaze, exp.data.pupil.processed.downsampled$Ygaze, pch=16, cex=.5, col=alpha(1), xpd=TRUE)
abline(h=768/2, v=1024/2, lty=1, col='white')
abline(h=768/2, v=1024/2, lty=2, col=1)
```

* Plot pupil size by subject
```{r}
par(cex=1.1, mar=c(5, 5, 2, 1), mgp=c(4,0.6,0))
bp <- sortBoxplot(exp.data.pupil.processed.downsampled$pupil.base.corrected.binned ~ exp.data.pupil.processed.downsampled$subject, col=alpha(1), decreasing=FALSE,
                  ylab="Pupil size", xlab="Subject", bty='n', pch=".", las=2)


```
#### data preparation for GAMM

* select interest period
```{r}
dat <- exp.data.pupil.processed.downsampled%>%
  filter(bin_low >= 4001 & bin_low <= 8001)%>% #select interest period
  rename(pupil_base="pupil.base.corrected.binned", time="bin_low")%>%
  arrange(subject,condition, agent, time, trial_number)%>%#order dataframe
  droplevels() %>% 
  mutate(test_event=ifelse(agent=="human"&condition=="new_goal", "human_new_goal",
  ifelse(agent=="human"&condition=="new_side", "human_new_side",
  ifelse(agent=="inanimate"&condition=="new_side", "inanimate_new_side",
  ifelse(agent=="inanimate"&condition=="new_goal", "inanimate_new_goal","")))))

dat$condition<-as.factor(dat$condition)
dat$agent<-as.factor(dat$agent)
dat$test_event<-as.factor(dat$test_event)

levels(dat$condition)
levels(dat$agent)
levels(dat$test_event)
summary(dat$test_event)
sum(is.na(dat$test_event))
dat$subject<-as.factor(dat$subject)
levels(dat$subject)
summary(dat)
```
#### fitting GAMM

```{r}
# Defining events (time series):
dat$Event <- interaction(dat$subject, dat$test_event, dat$trial_number, drop=TRUE)
levels(dat$Event)

library(mgcv)
gamm.int <- bam(pupil_base ~ condition*agent + 
           s(time, k=20) + 
           s(time, by= condition, k=20) +
             #argument by: fit different smooths for different levels of a categorical variables (factor-smooth interaction)
           s(time, by= agent, k=20) +
           s(Xgaze, Ygaze) +
           s(time, Event, bs='fs', m=1), 
           #bs="fs" is for random wiggly curves: produces a smooth curve for each level of a single factor, treating the curves as entirely random
           data=dat, discrete=TRUE, nthreads=40, method="ML") 

gamm.null <- bam(pupil_base ~ s(time, k=20)+ 
          + s(Xgaze, Ygaze)
          + s(time, Event, bs='fs', m=1)
          , data=dat, discrete=TRUE, nthreads=40, method="ML")

#full-null model comparison 
compareML(gamm.null, gamm.int) #AIC difference: 32.58, model gammm.int has lower AIC.

AIC(gamm.int, gamm.null)

summary(gamm.int)#linear interaction not significant, we remove it

gamm.full <- bam(pupil_base ~ condition + agent + 
           s(time, k=20) + 
           s(time, by= condition, k=20) +
           s(time, by= agent, k=20) +
           s(Xgaze, Ygaze) +
           s(time, Event, bs='fs', m=1), 
           data=dat, discrete=TRUE, nthreads=40, method="ML") 


library(mgcViz)
gam.check(gamm.full)


acf(resid(gamm.full), bty='n', main="ACF residuals model1")
acf(resid(gamm.full), plot=FALSE)
```


```{r}
#save gam_check
#remotes::install_github("m-clark/visibly")
library(visibly)
png("graphs/pupil_gamm.full_gam_check.png", units="cm", res=600, width=24, height = 18)
gam_check_plot<-plot_gam_check(gamm.full, scatter=TRUE)
print(gam_check_plot)
dev.off()

png("graphs/pupil_gamm.full_acf_plot.png", units="cm", res=600, width=12, height = 12)
acf_plot<-acf(resid(gamm.full), bty='n', main="")
print(acf_plot)
dev.off()
```
```{r}
summary(gamm.full)
###Fixed effects:
#in the new side condition, the pupil is in general larger than in the new goal condition, irrespective of the agent shown (t=2.36, P=0.019).
#Agent did not have a significant effect (t=0.47, P=0.638)
```

* difference curve
```{r}
#save data
plot_diff_gamm.full_newg_news<-plot_diff(gamm.full, view="time", 
           comp=list(condition=c("new_goal", "new_side")), rm.ranef=TRUE, 
          las=1, ylab="Est. difference in pupil size", 
          col=col2, hide.label = TRUE, plot = FALSE)

plot_diff_gamm.full_newg_news<-ggplot(data=plot_diff_gamm.full_newg_news, aes(x=time, y=est))+
  geom_hline(yintercept = 0)+
  geom_path(lty=2)+
  geom_ribbon(aes(x=time, ymin=est-CI, ymax=est+CI), alpha=0.2)+
  theme_bw()+
  scale_x_continuous(name="Time (in ms)", breaks=c(5000, 6000, 7000, 8000, 9000))+ ylab("Est. difference in pupil size")+ggtitle("New goal - New side")+
    theme(panel.grid.major.x = element_blank(),
        panel.grid.minor.x = element_blank(),
        panel.grid.major.y = element_blank(),
        panel.grid.minor.y = element_blank())

plot_diff_gamm.full_newg_news

ggsave(plot_diff_gamm.full_newg_news, filename = "graphs/pupilsize_diff_gamm.full_condition.png", width=8, height=6, scale=0.6, dpi=600)

```

* plotting partial effects
```{r}
pdf("graphs/pupil_gamm.full_partial_effects.pdf", width=10, height = 8)
plot(gamm.full, pages = 1, all.terms = TRUE, residuals = TRUE, rug=FALSE, pch = 1, cex = 0.5, seWithMean = TRUE, shade = TRUE, shade.col = "lightblue")
dev.off()

library(mgcViz)
b <- getViz(gamm.full)
png("graphs/pupil_gamm.full_partial_effect2.png", units="cm", res=600, width=24, height = 16)
p1<-plot(b, allTerms = T)+theme_classic() + labs(title = NULL)
print(p1, pages = 1 )
dev.off()

```
* Summed effects
```{r}

plot_smooth(gamm.full, view="time", cond=list(condition="new_goal"), rm.ranef=TRUE,
  v0=0, col=col2, lwd=2, lty=6, rug=FALSE, se=1.96,
  main="Estimated effects", ylab="Pupil (baselined)", las=1)

plot_smooth(gamm.full, view="time", cond=list(condition="new_side"), rm.ranef=TRUE,
  v0=0, col=col2, lwd=2, lty=6, rug=FALSE, se=1.96,
  main="Estimated effects", ylab="Pupil (baselined)", las=1)

```

```{r}
tmp <- gamm.full$model
plot_modelfit(gamm.full, view="time", event=tmp$Event,n = 2)
```
